### 2019-07-03 thesis-template/msc-thesis.tex
update

%In diesem Abschnitt sollen neben den Herausforderungen und der
%Problemstellung insbesondere die Ziele der Arbeit beschrieben werden.

\section{Structure of this Work}

@@

\section{GT Datasets / Methods}
\label{sec:gt}
\paragraph{Glyphs and Tensor Field Lines}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.4\textwidth]
    {img/tensorfieldlines.png}
  }
  \caption{tensor field lines: bold lines; ellipsoids: ellipsoid glyphs}
  \label{scheme}
\end{figure}\vskip 3pt
Glyphs are derived as a basis for the weighting profiles of the tensors in the following propagation scheme. They were chosen because it is an easy-to-use (understand) concept, which can also be interpreted nicely. In fact, these representations are used for the principal component analysis of direction vectors for the tensor field line extrapolation (integration) and also as transmission profiles for the light propagation scheme in a subsequent step. We can again emphasize glyphs as a crucial, central concept within this work. Tensor field lines (the type of tensor streamline analog that does not introduce artificial intertia) were implemented for ground truth (GT) data generation purposes. For example, when designing a FTLE-like field computation technique, we will need to know when tensor field lines diverge/converge.

@@

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/corner-TFL.png}
    \caption{``corner'' test field}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/spiral-TFL.png}
    \caption{``spiral'' test field}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/drain_old.png}
    \caption{``V'' test field    }
    \label{b)}
  \end{minipage}
%\caption{a) , b) , c) }
\label{test-fields}
\end{figure}

\textbf{Symmteric Tensor Fields}
We also aim to use a real data example of the $3D$ DTI-MRI diffusion tensor of the brain for a male subject at the age of 27 and another one featuring a human heart. As we visualize 2D symmetric second order tensor fields, we need to slice the dataset and crop the $2{\times}2$ matrices omitting the $z$-component of the original $3{\times}3$ matrices. Fig. \ref{brainglyphs} depicts the glyph representation of the test dataset for slice index $8$:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/brain_cyl_glyphs.png}
    \label{a)}
    \caption*{cylinder glyphs}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/brain_ellipsoid_glyphs.png}
    \label{b)}
    \caption*{ellipsoid glyphs}
  \end{minipage}
\caption{``brain''-test field}
\label{brainglyphs}
\end{figure}

@@

\begin{figure}[!t]
\centering
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/drain(alt).png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/drain(alt)-TFL.png}
    \label{b)}
  \end{minipage}
\caption{a) ``drain''-testfield, b) tensor field lines for a)}
\label{drain}
\end{figure}

@@

\begin{figure}[!t]
\centering
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/rings.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/tensorfieldlines.png}
    \label{b)}
  \end{minipage}
\caption{a) ``rings'' test field, b) tensor field lines for a)}
\label{rings}
\end{figure}
The ``rings'' test field, depicted in Fig. \ref{rings} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/spiral.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/spiral-TFL.png}
    \label{b)}
  \end{minipage}
\caption{a) ``spiral''-test field, b) tensor field lines for a)}
\label{spiral}
\end{figure}
The ``spiral'' test field, depicted in Fig. \ref{spiral} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/inverse.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/inverse-TFL.png}
    \label{b)}
  \end{minipage}
\caption{a) ``inverse''-test field, b) tensor field lines for a)}
\label{inverse}
\end{figure}
With the \enquote{inverse} test field we want to demonstrate the directive separation of intensities into four quadrants.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/corner.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/corner-TFL.png}
    \label{b)}
  \end{minipage}
\caption{a) ``corner''-test field, b) tensor field lines for a)}
\label{corner}
\end{figure}
The test field ``corner'', depicted in Fig \ref{corner}, constitutes a sustained turn over the grid representing the redirection or forwarding along domain corners. To have some real examples involved, we apply the invented technique to both DTI-MRI \enquote{brain} and \enquote{heart} scans, which can be observed in Fig. \ref{real1}.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/brain_ellipsoid_glyphs.png}
    \label{a)}
    \caption*{sphere glyphs}
  \end{minipage}
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/brain_cyl_glyphs.png}
    \label{b)}
    \caption*{cyclinder glyphs}
  \end{minipage}
\caption{Real data examples: \enquote{brain} dataset}
\label{real1}
\end{figure}
Observing the \enquote{brain} dataset in Fig \ref{real1}, one may notice that it introduces a lot of variances to tensor magnitude and linear anisotropy. We choose this DTI-MRI scan as an example-par-excellence for real application data given in a reasonable resolution.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.35\textwidth}
    \includegraphics[width=\textwidth]{img/heart_org.png}
    \label{a)}
    \caption*{tensor magnitude}
  \end{minipage}
  \begin{minipage}{0.35\textwidth}
      \includegraphics[width=\textwidth]{img/heart_cyl_glyphs.png}
    \label{b)}
    \caption*{cyclinder glyphs}
  \end{minipage}
\caption{Real data examples \enquote{heart} dataset}
\label{real2}
\end{figure}
When we look at this DT-MRI scan of a human heart we can notice the elongated shape, which we consider to represent the aorta and a part of the rest of the heart. If we consider the glyph representation, we again infer, that we incorporate lots of anisotropic regions with varying tensor magnitudes here.

@@

\paragraph{Global Illumination Methods}
Since we use a specific category of global illumination methods as a basis, we decide to only discuss lattice based methods which work, corresponding to their name, on a grid. Remark, that Global Illumination methods classically focus on realistic indirect lighting. Thus, they aim on producing a most realistic image, what in our case turns out to be a most informative image in correspondence. The most early work in this field is propably the Discrete Ordinates Method (DOM) by Chandrasekhar \cite{Chandras}, which is used specifically for computing radiative transfer in participating media (volume rendering). The radiative transfer equation is discretized spatially and in the angular domain. The radiance distribution (PDF) is stored in a grid and light is distributed and exchanged between neighboring volume elements which reduces the overall effort to local and independent computations. Furthermore, there has been more recent work on describing the light propagation as a diffusion process within a simple photon transport model based on the Lattice-Boltzmann method \cite{geist}. DOMs were recently revisited by Fattal \cite{fattal}. This work adresses artifacts stemming from spatial discretization, which are called \enquote{ray effects}, and from successive interpolation called \enquote{light smearing}. These methods are considered to be too slow to \enquote{run at interactive framerates} by Dachsbacher et al. \cite{dachsbacher}, though \enquote{efficient and potentially highly parallel}. Our work is inspired by Lattice-Boltzmann methods (volume grid radiosity) and the lattice-based discrete ordinate methods concerning the light propagation scheme.

@@

\section{Light Propagation Scheme Behaviour}
To evaluate the plausibility of the light distribution results obtained from the light propagation scheme, we plot the transmission profiles overlaid with the resulting intensity profiles for several synthetic test fields, whereas red: tensor glyphs (transmission profiles) and blue: intensity profiles are both depicted in polar coordinates:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/isotropic.png}
    \label{a)}
    \caption*{light src in top center}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/isotropic-all-center.png}
    \label{b)}
    \caption*{light src in center}
  \end{minipage}
\caption{final light propagation distributions for ``isotropy''-testfield}
\label{isotropic}
\end{figure}
One can observe here, how light is propagated with no directional bias leading to an absolutely isotropic distribution in case  b), which is symmetric as expected. This result is considered to be the absolute ground truth example for the light propagation scheme, which is evaluated separately in sect. \ref{sec:att} and \ref{sect:conserve}. We can also depict the final distribution as a scalar field when using the total energy in each cell as scalar:

@@

    \includegraphics[width=\textwidth]{img/iso_scalar.png}

@@

    \includegraphics[width=\textwidth]{img/iso_sat.png}

@@

    \includegraphics[width=\textwidth]{img/rings-two-special1.png}

@@

    \includegraphics[width=\textwidth]{img/rings-two-special2.png}

@@

    \includegraphics[width=\textwidth]{img/cos_parallel.png}

@@

    \includegraphics[width=\textwidth]{img/cos_anti.png}

@@

    \includegraphics[width=\textwidth]{img/spiral-two-dense.png}

@@

    \includegraphics[width=\textwidth]{img/spiral-two-wide.png}

@@

    \includegraphics[width=\textwidth]{img/spiral-temp2.png}

@@

    \includegraphics[width=\textwidth]{img/spiral_full.png}

@@

\begin{figure}[t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/isotropic2.png}
    \label{a)}
    \caption*{a) Glyphs}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/iso_ftle.png}
    \label{b)}
    \caption*{b) LTG($0^\circ$) for a)}
  \end{minipage}
  \caption{\enquote{isotropic} test field}
\label{ftle_iso}
\end{figure}

@@

\begin{figure}[t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/iso_volume.png}
    \label{a)}
    \caption*{a) Volume}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/iso_isovolume.png}
    \label{b)}
    \caption*{b) IsoVolume Filter for a)}
  \end{minipage}
  \caption{\enquote{isotropic} test field}
\label{ftle_iso}
\end{figure}

Now we will discuss a synthetic example designed to evaluate the performance of the FTLE. As a little warmup, we will have an experiment exhibting bare, raw tensor field lines diverging w.r.t. the $y$-axis:

\begin{figure}[t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/drain(alt)-TFL.png}
    \label{a)}
    \caption*{a) TFLs and glyphs}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_drain_alt.png}
    \label{b)}
    \caption*{b) LTG($90^\circ$) for a)}
  \end{minipage}
  \caption{``drain'' test field}
\label{ftle_base}
\end{figure}
What we can observe here, is a feature called a ``ridge'' occuring as a high value line segment (red vertical track) which is already quite obvious for human vision. A gradient filter could be applied to extract and segment edges from the image as a preprocessing step for computer vision, which would look like the image in Fig. \ref{grad)}:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_drain_alt.png}
    \label{grada)}
    \caption*{a) LTG($90^\circ$) }
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/drain_alt_gradient.png}
    \label{gradb)}
    \caption*{b) gradient for a)}
  \end{minipage}
  \caption{``drain'' test field}
  \label{grad)}
\end{figure}

 These ridges represent ROIs with underlying Lagrangian Coherent Structures, which we already discussed in sect. \ref{sect:motiv}. It could for example represent a stress field inside of a mechanical tool with a central axis with maximum stress condensation. 
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/drain_top.PNG}\vskip 10pt
    \includegraphics[width=\textwidth]{img/drain_alt-profile.PNG}
    \caption*{Cross section on LTG volume representation}	
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/drain_alt_isovolume.PNG}
     \caption*{IsoVolume filter on }
    \label{b)}
  \end{minipage}
  \caption{LTG volume representation: \enquote{drain} test field}
\label{drain_contour}
\end{figure}
For the ``drain'' test field we consider the central channel which produces a clear line in the cross section. The light transport gradient responds most inside of the channel region and on the top T-arms independently of the stimulus direction. Next, we will analyze the \enquote{inverse} test field in similar fashion:

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/inverse-TFL.png}
    \label{a)}
    \caption*{a) TFLs and glyphs }
  \end{minipage}
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/inverse-FTLE-raw.png}
    \label{b)}
    \caption*{b) LTG($90^\circ$) for a)}
  \end{minipage}
  \caption{``inverse''-test field}
\label{inverse-ftle}
\end{figure}

We also evaluated the approach for the ``inverse'' test field, whereas we can observe height ridges for the $90^\circ$ slice in any position with diverging tensor field lines w.r.t. the light source direction. This is certainly not the case for the upper (missing) ridge whereas tensor field lines converge with respect to the UP ($90^\circ$) direction. If we determine the ridges pointing into the diverging direction, this proves that the method is capable of detecting ridges from parallel up to orthogonal to the light source direction, but not in antiparallel directions, which exhibit converging tensor field lines as bidirectional counterpart. Luckily, we measured the gradient for every possible sampled light source direction which complements this information into another slice (and the resulting volume of stacked slices).

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/inverse_isovolume.png}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/inverse_surface.png}
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/inverse_contour.png}
  \end{minipage}
    \caption{IsoVolume Filter on LTG volume representation: ``inverse'' test field}
\label{inverse_contour}
\end{figure}
Note that the $z$-axis represents the light source direction angle $\omega$ in this case. The contour filter reveals the star-shaped structure of the tensor field which rotates along the angle axis. This is the effect we already captured for the single slice: For each axis direction we can detect ridges from parallel up to orthogonal to our predefined stimulus (light source) direction, but not in antiparallel direction. This causes the star shape to miss one branch or ray which is in turn the antiparallel direction missing for each light source direction, emphasizing the impression of rotation. We can also try to segment height ridges represented as surfaces from the resulting volume grids as depicted in Fig \ref{inverse_contour}. The orthogonal ridges in this example reach out to the edges because they are gaining a bias by never offsetting the light source in a transmission direction (ellipsoid orientation) for orthogonals. This turns out to be a general feature of the method, as we will later verify in Fig. \ref{res_ftle_volume} and \ref{star_ftle}.

>>>>

%In diesem Abschnitt sollen neben den Herausforderungen und der
%Problemstellung insbesondere die Ziele der Arbeit beschrieben werden. 
\section{Structure of this Work}

@@

\section{GT Datasets / Methods}
\label{sec:gt}
\paragraph{Glyphs and Tensor Field Lines}
%\begin{figure}[!t]
%  \centering
% {
%    \includegraphics[width=0.4\textwidth]
%    {img/tensorfieldlines.png}
%  }
%  \caption{tensor field lines: bold lines; ellipsoids: ellipsoid glyphs}
%  \label{scheme}
%\end{figure}
Glyphs are derived as a basis for the weighting profiles of the tensors in the following propagation scheme. They were chosen because it is an easy-to-use (understand) concept, which can also be interpreted nicely. In fact, these representations are used for the principal component analysis of direction vectors for the tensor field line extrapolation (integration) and also as transmission profiles for the light propagation scheme in a subsequent step. We can again emphasize glyphs as a crucial, central concept within this work. Tensor field lines (the type of tensor streamline analog that does not introduce artificial intertia) were implemented for ground truth (GT) data generation purposes. For example, when designing a FTLE-like field computation technique, we will need to know when tensor field lines diverge/converge. Examples for glyphs (ellipsoids) and tensor field lines (thin lines) are depicted in Fig. \ref{test-fields}.

@@

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/corner-TFL.png}
    \caption*{``corner''}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/spiral-TFL.png}
    \caption*{``spiral''}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/drain_old.png}
    \caption*{``V''}
    \label{b)}
  \end{minipage}
\caption{Named test fields}
\label{test-fields}
\end{figure}

@@

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.8\textwidth]{img/drain(alt).png}
    \caption*{a) glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/drain(alt)-TFL.png}
    \caption*{b) tensor field lines for a)}
    \label{b)}
  \end{minipage}
\caption{``drain'' test field}
\label{drain}
\end{figure}

@@

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.8\textwidth]{img/rings.png}
    \caption*{a) glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
   \centering
    \includegraphics[width=0.8\textwidth]{img/tensorfieldlines.png}
	\caption*{b) tensor field lines for a)}
    \label{b)}
  \end{minipage}
\caption{``rings'' test field}
\label{rings}
\end{figure}
The ``rings'' test field, depicted in Fig. \ref{rings} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/spiral.png}
    \caption*{a) glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/spiral-TFL.png}
    \caption*{b) tensor field lines for a)}
    \label{b)}
  \end{minipage}
\caption{``spiral''-test field}
\label{spiral}
\end{figure}
The ``spiral'' test field, depicted in Fig. \ref{spiral} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/inverse.png}
    \caption*{a) glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/inverse-TFL.png}
    \caption*{b) tensor field lines for a)}
    \label{b)}
  \end{minipage}
\caption{``inverse'' test field}
\label{inverse}
\end{figure}
With the \enquote{inverse} test field we want to demonstrate the directive separation of intensities into four quadrants.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/gyre.png}
    \caption*{a) glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/gyre-TFL.png}
    \caption*{b) tensor field lines for a)}
    \label{b)}
  \end{minipage}
\caption{a) ``gyre''-test field, b) tensor field lines for a)}
\label{gyre}
\end{figure}
The test field ``gyre'', depicted in Fig \ref{gyre}, constitutes a sustained turn over the grid representing the redirection or forwarding along domain corners.

\textbf{Symmteric Tensor Fields}
We also aim to use a real data example of the $3D$ DTI-MRI diffusion tensor of the brain for a male subject at the age of 27 and another one featuring a human heart. As we visualize 2D symmetric second order tensor fields, we need to slice the dataset and crop the $2{\times}2$ matrices omitting the $z$-component of the original $3{\times}3$ matrices. Fig. \ref{brainglyphs} depicts the glyph representation of the test dataset for slice index $15$. To have some real examples involved, we apply the invented technique to both DTI-MRI \enquote{brain} and \enquote{heart} scans, which can be observed in Fig. \ref{real1} and \ref{real2}.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.8\textwidth]{img/brain_org.png}
    \label{a)}
    \caption*{tensor magnitude}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.8\textwidth]{img/brain_cyl_glyphs.png}
    \label{b)}
    \caption*{cyclinder glyphs}
  \end{minipage}
\caption{Real data examples: \enquote{brain} dataset}
\label{real1}
\end{figure}
Observing the \enquote{brain} dataset in Fig \ref{real1}, one may notice that it introduces a lot of variances to tensor magnitude and linear anisotropy. We choose this DTI-MRI scan as an example-par-excellence for real application data given in a reasonable resolution.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/heart_org.png}
    \label{a)}
    \caption*{tensor magnitude}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
      \includegraphics[width=0.8\textwidth]{img/heart_cyl_glyphs.png}
    \label{b)}
    \caption*{cyclinder glyphs}
  \end{minipage}
\caption{Real data examples \enquote{heart} dataset}
\label{real2}
\end{figure}
When we look at this DT-MRI scan of a human heart we can notice the elongated shape, which we consider to represent the aorta and a part of the rest of the heart. If we consider the glyph representation, we again infer, that we incorporate lots of anisotropic regions with varying tensor magnitudes here.

@@

\paragraph{Global Illumination Methods}
Since we use a specific category of global illumination methods as a basis, we decide to only discuss lattice based methods which work, corresponding to their name, on a grid. Remark, that Global Illumination methods classically focus on realistic indirect lighting. Thus, they aim on producing a most realistic image, what in our case turns out to be a most informative image in correspondence. The most early work in this field is propably the Discrete Ordinates Method (DOM) by Chandrasekhar \cite{Chandrasekhar}, which is used specifically for computing radiative transfer in participating media (volume rendering). The radiative transfer equation is discretized spatially and in the angular domain. The radiance distribution (PDF) is stored in a grid and light is distributed and exchanged between neighboring volume elements which reduces the overall effort to local and independent computations. Furthermore, there has been more recent work on describing the light propagation as a diffusion process within a simple photon transport model based on the Lattice-Boltzmann method \cite{geist}. DOMs were recently revisited by Fattal \cite{fattal}. This work adresses artifacts stemming from spatial discretization, which are called \enquote{ray effects}, and from successive interpolation called \enquote{light smearing}. These methods are considered to be too slow to \enquote{run at interactive framerates} by Dachsbacher et al. \cite{dachsbacher}, though \enquote{efficient and potentially highly parallel}. Our work is inspired by Lattice-Boltzmann methods (volume grid radiosity) and the lattice-based discrete ordinate methods concerning the light propagation scheme.

@@

\section{Light Propagation Scheme Behaviour}
To evaluate the plausibility of the light distribution results obtained from the light propagation scheme, we plot the transmission profiles overlaid with the resulting intensity profiles for several synthetic test fields, whereas red: tensor glyphs (transmission profiles) and blue: intensity profiles are both depicted in polar coordinates:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.8\textwidth]{img/isotropic.png}
    \label{a)}
    \caption*{light src in top center}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.8\textwidth]{img/isotropic-all-center.png}
    \label{b)}
    \caption*{light src in center}
  \end{minipage}
\caption{final light propagation distributions for ``isotropy''-testfield}
\label{isotropic}
\end{figure}
One can observe here, how light is propagated with no directional bias leading to an absolutely isotropic distribution in case  b), which is symmetric as expected. This result is considered to be the absolute ground truth example for the light propagation scheme, which is evaluated separately in sect. \ref{sec:att} and \ref{sect:conserve}. We can also depict the final distribution as a scalar field when using the total energy in each cell as scalar:

@@

    \includegraphics[width=0.9\textwidth]{img/iso_scalar.png}

@@

    \includegraphics[width=0.9\textwidth]{img/iso_sat.png}

@@

    \includegraphics[width=0.9\textwidth]{img/rings-two-special1.png}

@@

    \includegraphics[width=0.9\textwidth]{img/rings-two-special2.png}

@@

    \includegraphics[width=0.9\textwidth]{img/cos_parallel.png}

@@

    \includegraphics[width=0.9\textwidth]{img/cos_anti.png}

@@

    \includegraphics[width=0.9\textwidth]{img/spiral-two-dense.png}

@@

    \includegraphics[width=0.9\textwidth]{img/spiral-two-wide.png}

@@

    \includegraphics[width=0.9\textwidth]{img/spiral-temp2.png}

@@

    \includegraphics[width=0.9\textwidth]{img/spiral_full.png}

@@

\begin{figure}[t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[height=\textwidth]{img/isotropic2.png}
    \label{a)}
    \caption*{a) Glyphs}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[height=\textwidth]{img/iso_ftle.png}
    \label{b)}
    \caption*{b) LTG($0^\circ$) for a)}
  \end{minipage}
  \caption{\enquote{isotropic} test field}
\label{ftle_iso}
\end{figure}

@@

\begin{figure}[t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/iso_volume.png}
    \label{a)}
    \caption*{a) Volume}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.8\textwidth]{img/iso_isovolume.png}
    \label{b)}
    \caption*{b) IsoVolume Filter for a)}
  \end{minipage}
  \caption{\enquote{isotropic} test field}
\label{ftle_iso}
\end{figure}
To verify the concept of absorption implicitly, we incorporate absorption into test field \enquote{absorb}, whereas all cells are absorbing energy except the vertical center line, which is assumed to produce a clear ridge. 
\begin{figure}[t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/absorb.png}
    \label{a)}
    \caption*{a) Glyphs}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.9\textwidth]{img/absorb_ftle.png}
    \label{b)}
    \caption*{b) LTG($90^\circ$) for a)}
  \end{minipage}
  \caption{\enquote{absorb} test field}
\label{ftle_absorb}
\end{figure}

We observe the clear straight ridge in the example in Fig. \ref{ftle_absorb}. As expected, the ridge is located on the vertivcal center line. Note, that we (min-max) normalize each transmission profile with the highest mean transmission rate (tensor magnitude) in the field. From this test case, we can conclude, that regions with a comparably small tensor magnitude would not obtain any FTLE field response, which we will later observe for the real datasets \enquote{brain} and \enquote{heart} exhibiting such areas. We will now discuss synthetic examples designed to evaluate the performance of the FTLE. As a little warmup, we will have an experiment exhibting bare, raw tensor field lines diverging w.r.t. the $y$-axis:

\begin{figure}[t]
\centering
  \begin{minipage}{0.3\textwidth}
\centering
    \includegraphics[width=\textwidth]{img/drain(alt)-TFL.png}
    \label{a)}
    \caption*{a) TFLs and glyphs}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_drain_alt.png}
    \label{b)}
    \caption*{b) LTG($90^\circ$) for a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/drain_alt_gradient.png}
    \label{b)}
    \caption*{c) gradient for b)}
  \end{minipage}
  \caption{``drain'' test field}
\label{ftle_base}
\end{figure}
What we can observe here, is a feature called a ``ridge'' occuring as a high value line segment (red vertical track) which is already quite obvious for human vision. A gradient filter could be applied to extract and segment edges from the image as a preprocessing step for computer vision, which would look like the image in Fig. \ref{ftle_base} c).

These ridges represent ROIs with underlying Lagrangian Coherent Structures, which we already discussed in sect. \ref{sect:motiv}. It could for example represent a stress field inside of a mechanical tool with a central axis with maximum stress condensation. 
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
  \includegraphics[width=0.9\textwidth]{img/drain_top.PNG}\vskip 10pt
    \includegraphics[width=0.9\textwidth]{img/drain_alt-profile.PNG}
    \caption*{Cross section on LTG volume}	
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/drain_alt_isovolume.PNG}
     \caption*{IsoVolume Filter}
    \label{b)}
  \end{minipage}
  \caption{LTG volume representation: \enquote{drain} test field}
\label{drain_contour}
\end{figure}
For the ``drain'' test field we consider the central channel which produces a clear line in the cross section. The light transport gradient responds most inside of the channel region and on the top T-arms independently of the stimulus direction. For the sake of completeness, we show that the approach's functionality also for diagonal directions in Fig. \ref{diag-ftle}.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/diagDrain.png}
    \label{a)}
    \caption*{a) TFLs and glyphs}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/diagDrain-ftle.png}
    \label{b)}
    \caption*{b) LTG($225^\circ$) for a)}
  \end{minipage}
  \caption{``diagonal drain'' test field}
\label{diag-ftle}
\end{figure}

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/inverse-TFL.png}
    \label{a)}
    \caption*{a) TFLs and glyphs }
  \end{minipage}
  \begin{minipage}[!t]{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/inverse-FTLE-raw.png}
    \label{b)}
    \caption*{b) LTG($90^\circ$) for a)}
  \end{minipage}
  \caption{``inverse''-test field}
\label{inverse-ftle}
\end{figure}

We also evaluated the approach for the ``inverse'' test field, depicted in Fig. \ref{inverse-ftle}, whereas we can observe height ridges for the $90^\circ$ slice in any position with diverging tensor field lines w.r.t. the light source direction. This is certainly not the case for the upper (missing) ridge whereas tensor field lines converge with respect to the UP ($90^\circ$) direction. If we determine the ridges pointing into the diverging direction, this proves that the method is capable of detecting ridges from parallel up to orthogonal to the light source direction, but not in antiparallel directions, which exhibit converging tensor field lines as bidirectional counterpart. Luckily, we measured the gradient for every possible sampled light source direction which complements this information into another slice (and the resulting volume of stacked slices).

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/inverse_isovolume.png}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/inverse_surface.png}
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/inverse_contour.png}
  \end{minipage}
    \caption{IsoVolume Filter on LTG volume representation: ``inverse'' test field}
\label{inverse_contour}
\end{figure}
Note that the $z$-axis represents the light source direction angle $\omega$ in this case. The contour filter reveals the star-shaped structure of the tensor field which rotates along the angle axis. This is the effect we already captured for the single slice: For each axis direction we can detect ridges from parallel up to orthogonal to our predefined stimulus (light source) direction, but not in antiparallel direction. This causes the star shape to miss one branch or ray which is in turn the antiparallel direction missing for each light source direction, emphasizing the impression of rotation. We can also try to segment height ridges represented as surfaces from the resulting volume grids as depicted in Fig \ref{inverse_contour}. The orthogonal ridges in this example reach out to the edges because they are gaining a bias by never offsetting the light source in a transmission direction (ellipsoid orientation) for orthogonals. This turns out to be a general feature of the method, as we will later verify in Fig. \ref{res_ftle_volume} and \ref{star-ftle}.
